Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-14 paper/manuscript.Rmd
edit to P1
infer swim

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between two small islands (Figure
1, see Figure 2 for location), a distance of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether
swimming behaviour was common among GPS radio-collared caribou in the population
[n = 29, for collaring details see, @Peignier2019]. We identified swimming
events as two consecutive GPS locations (2-hour relocation rates)  from an
individual that occurred on different islands. Northeastern Newfoundland typically
experiences pack ice during winter and caribou are known locally to travel
between islands by walking over the ice. We restricted our GPS data to the
ice-free period of the year (April 1 to December 31) and identified `r nevents`
swimming events over three years (Figure 2). In addition to our own visual and
remotely sensed observations, residents of Fogo Island have also reported regularly
observing caribou swimming between islands.

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. On average, caribou swam approximately `r neventidMean`
(range = `r neventidMin`-`r neventidMax`) times per year and remained on each
island for approximately `r dur` (range = `r durMin`-`r durMax`) days before
swimming again. Based on swimming data, we separated the population into three
apparent groups of individuals: those that swam regularly between islands (n =
3, every ~`r indMean` days); those that rarely engaged in swimming events (n =
9, every ~ `r ind2Mean` days); and those that never swam (n = 17).

>>>>

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between two small islands (Figure
1, see Figure 2 for location), a distance of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether
swimming behaviour was common among GPS radio-collared caribou in the population
[n = 29, for collaring details see, @Peignier2019]. We identified an
individual's swimming events as two consecutive GPS locations (2-hour relocation
rates)  on different islands. Northeastern Newfoundland typically experiences
pack ice during winter and caribou are known locally to travel between islands
by walking over the ice. We restricted our GPS data to the ice-free period of
the year (April 1 to December 31) and identified `r nevents` swimming events
over three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported regularly observing
caribou swimming between islands.

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Of the caribou that swam, caribou swam on average approximately `r neventidMean`
(range = `r neventidMin`-`r neventidMax`) times per year and remained on each
island for approximately `r dur` (range = `r durMin`-`r durMax`) days before
swimming again. Based on GPS data to infer swimming, we separated the population into three
apparent groups of individuals: those that swam regularly between islands (n =
3, every ~`r indMean` days); those that rarely engaged in swimming events (n =
9, every ~ `r ind2Mean` days); and those that never swam (n = 17).
